---
title: "Risk of Metastasis/Death by Baseline Risk"
subtitle: "Analysis"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Analytic Data Import

```{r import}
# load libraries
library(tidyverse)
library(here)
library(survival)
library(gt)
library(gtsummary)
set.seed(86453768)

# Data file path
# TODO: Fill in your own file path here
path_data <- file.path("O:\\Outcomes\\Andrew\\Analytic Projects\\1 Active\\Vickers Relative risk of PCa Death by Baseline Risk (SPCG4 Cohort)\\2020-06-01")

# importing master analytic data set
df_master <-
  readRDS(
    file = here(path_data, "df_master.rds")
  )

```

# Set up Framework for Analysis

```{r analysisetup}

# Create a dataset with four different groupings - grouped by quartiles of predicted probability of BCR score
df_analysis <-
  df_master %>%
  nest(df_model = -c(quartile)) %>%
  arrange(quartile) %>%
  filter(!is.na(quartile)) %>%
  mutate(
    # Save out ranges of predicted probability values for each
    lower_pr = map_dbl(df_model, ~ min(..1$preopbcr_pr5, na.rm = TRUE)),
    upper_pr = map_dbl(df_model, ~ max(..1$preopbcr_pr5, na.rm = TRUE))
  )

```

# Summary Tables and Figures

```{r table1}

# Table 1 for all who aren't missing data
gt_summary_all <-
  df_master %>%
  select(age, psa, bxggs, tstage, detectionmode, preopbcr_pr5, rp) %>%
  tbl_summary(
    digits = list("age" ~ 0), # "psa" ~ c(0, 1, 1)),
    label = list("preopbcr_pr5" ~ "Risk of BCR at 5 Years",
                 "rp" ~ "Immediate RP")
  ) %>%
  modify_footnote(everything() ~ NA)

gt_summary_all

# Table 1 by quartiles
gt_summary_quartiles <-
  df_master %>%
  group_by(quartile) %>%
  mutate(
    quartilelabel =
      case_when(
        quartile == 1 ~ "First",
        quartile == 2 ~ "Second",
        quartile == 3 ~ "Third",
        quartile == 4 ~ "Fourth"
      ),
    rangemin = min(preopbcr_pr5, na.rm = TRUE),
    rangemax = max(preopbcr_pr5, na.rm = TRUE),
    label = as.character(str_glue("{quartilelabel} Quartile:<br> Risk {style_percent(rangemin)}% - {style_percent(rangemax)}%")) %>% factor()
  ) %>%
  ungroup() %>%
  select(age, psa, bxggs, tstage, detectionmode, preopbcr_pr5, rp, label) %>%
  tbl_summary(
    by = label,
    digits = list("age" ~ 0), #, "psa" ~ c(0, 1, 1)),
    label = list("preopbcr_pr5" ~ "Risk of BCR at 5 Years",
                 "rp" ~ "Radical Prostatectomy")
  ) %>%
  modify_footnote(everything() ~ NA)

gt_summary_quartiles

# Histogram of risk
gg_risk_histogram <-
  ggplot(
    df_master,
    aes(x = preopbcr_pr5)
  ) +
  geom_histogram(bins = 35) +
  labs(
    y = "",
    x = "Risk of BCR within 5 years"
  )

gg_risk_histogram

# Save out counts

# Number with metastasis
Nmets <- nrow(df_master %>% filter(mets == 1))
Nmetscensored <- nrow(df_master %>% filter(mets == 1 & dead == 0 & dod == 0))
Nmetsdead <- nrow(df_master %>% filter(mets == 1 & dead == 1))
Nmetsdod <- nrow(df_master %>% filter(mets == 1 & dead == 1 & dod == 1))

# Save out median followup (with quartiles) for survivors
medianfu <-
  df_master %>%
  filter(dead == 0) %>%
  {quantile(.$ttdead, c(0.25, 0.5, 0.75))}

```

# Distant Metastasis or Prostate Cancer Death by Baseline Risk Group

This analysis is performed for both outcomes (prostate cancer metastasis and prostate cancer death). A separate competing risks regression model is created in each quartile of patients with the covariate of interest being treatment group (immediate RP or delayed RP) so that the estimate in each group can be meta-analyzed later. As a sensitivity analysis, these analyses were repeating adjusting for age, as the version of the MSK BCR model used to calculate risk of BCR for this analysis did not include age as a covariate.

## Create Competing Risk Models

```{r crranalysis}

# Create Cox model for each cohort separately (exclude NA)
df_analysis_crr <-
  # Save out rows for mets outcome and rows for DOD outcome
  bind_rows(
    df_analysis %>% mutate(outcome = "mets"),
    df_analysis %>% mutate(outcome = "dod")
  ) %>%
  # Rename outcome variables to be the same in each dataset for consistency when creating models
  mutate(
    df_model =
      map2(
        df_model, outcome,
        ~ ..1 %>%
          rename(
            ttout := as.character(str_glue("tt{..2}")),
            out := as.character(str_glue("{..2}")),
            crr_outcome := as.character(str_glue("crr_{..2}_outcome")))
      ),
    # Create competing risk model for each outcome
    model_crr =
      map2(
        df_model, outcome,
        ~ cmprsk::crr(
          ftime = ..1[["ttout"]],
          fstatus = ..1[["crr_outcome"]],
          cov1 = ..1[c("rp")],
          failcode = 1, # Event (mets or DOD)
          cencode = 0 # Censored
        )
      ),
    
    # Sensitivity analysis adjusting for age
    model_age_crr =
      map2(
        df_model, outcome,
        ~ cmprsk::crr(
          ftime = ..1[["ttout"]],
          fstatus = ..1[["crr_outcome"]],
          cov1 = ..1[c("rp", "age")],
          failcode = 1, # Event (mets or DOD)
          cencode = 0 # Censored
        )
      ),
    
    
    # Pull out SHR, CI, p value
    model_results =
      pmap(
        list(model_crr, lower_pr, upper_pr),
        # First, get SHR and 95% CI
        ~ summary(..1)$conf.int %>%
          as_tibble() %>%
          set_names(c("shr", "shr_rev", "shr_lb", "shr_ub")) %>%
          # Add in coefficient, SE, p value
          bind_cols(
            summary(..1)$coef %>%
              as_tibble() %>%
              set_names(c("coef", "expcoef", "se_coef", "z", "p.value")) %>%
              select(-expcoef, -z)
          ) %>%
          # Add labels
          bind_cols(
            tribble(
              ~label,
              str_glue("Risk {style_percent(..2)}% - {style_percent(..3)}%")
            )
          )
      ),
    # Pull out results for sensitivity analysis model adjusted for age
    model_age_results =
      pmap(
        list(model_age_crr, lower_pr, upper_pr),
        # First, get SHR and 95% CI
        ~ summary(..1)$conf.int %>%
          as_tibble(rownames = "var") %>%
          set_names(c("var", "shr", "shr_rev", "shr_lb", "shr_ub")) %>%
          # Add in coefficient, SE, p value
          left_join(
            summary(..1)$coef %>%
              as_tibble(rownames = "var") %>%
              set_names(c("var", "coef", "expcoef", "se_coef", "z", "p.value")) %>%
              select(-expcoef, -z),
            by = "var"
          ) %>%
          # Keep only RP
          filter(var == "rp") %>%
          # Add labels
          bind_cols(
            tribble(
              ~label,
              str_glue("Risk {style_percent(..2)}% - {style_percent(..3)}%")
            )
          )
      )
  )

# Create table of CRR results
gt_model_crr <-
  df_analysis_crr %>%
  select(outcome, model_results) %>%
  unnest(model_results) %>%
  gt() %>%
  cols_move_to_start(columns = "label") %>%
  cols_hide(columns = c("shr_rev", "coef", "se_coef", "p.value", "outcome")) %>%
  fmt(columns = c("shr", "shr_lb", "shr_ub"), fns = style_ratio) %>%
  cols_label(
    label = "Risk Score",
    shr = "SHR",
    shr_lb = "95% CI",
  ) %>%
  tab_row_group(
    label = "Prostate Cancer Death",
    rows = outcome == "dod"
  ) %>%
  cols_merge(columns = c("shr_lb", "shr_ub"), pattern = "{1}, {2}") %>%
  tab_row_group(
    label = "Distant Metastasis",
    rows = outcome == "mets"
  )

gt_model_crr

# Create same table for sensitivity analysis with age
gt_model_age_crr <- 
  df_analysis_crr %>%
  select(outcome, model_age_results) %>%
  unnest(model_age_results) %>%
  gt() %>%
  cols_move_to_start(columns = "label") %>%
  cols_hide(columns = c("var", "shr_rev", "coef", "se_coef", "p.value", "outcome")) %>%
  fmt(columns = c("shr", "shr_lb", "shr_ub"), fns = style_ratio) %>%
  cols_label(
    label = "Risk Score",
    shr = "SHR",
    shr_lb = "95% CI",
  ) %>%
  tab_row_group(
    label = "Prostate Cancer Death",
    rows = outcome == "dod"
  ) %>%
  cols_merge(columns = c("shr_lb", "shr_ub"), pattern = "{1}, {2}") %>%
  tab_row_group(
    label = "Distant Metastasis",
    rows = outcome == "mets"
  )

gt_model_age_crr

# Create table with both
gt_model_full_crr <-
  df_analysis_crr %>%
  select(outcome, model_results, model_age_results) %>%
  mutate(model_age_results = map(model_age_results, ~ .x %>% rename_with(~str_glue("{.}_age")))) %>%
  unnest(model_results) %>%
  unnest(model_age_results) %>%
  gt() %>%
  cols_move_to_start(columns = "label") %>%
  cols_hide(columns = c("shr_rev", "coef", "se_coef", "p.value", "outcome", "label_age",
                        "var_age", "shr_rev_age", "coef_age", "se_coef_age", "p.value_age")) %>%
  fmt(columns = c("shr", "shr_lb", "shr_ub", "shr_age", "shr_lb_age", "shr_ub_age"), fns = style_ratio) %>%
  cols_label(
    label = "Risk Score",
    shr = "SHR",
    shr_lb = "95% CI",
    shr_age = "SHR",
    shr_lb_age = "95% CI"
  ) %>%
  cols_merge(columns = c("shr_lb", "shr_ub"), pattern = "{1}, {2}") %>%
  cols_merge(columns = c("shr_lb_age", "shr_ub_age"), pattern = "{1}, {2}") %>%
  tab_row_group(
    label = "Prostate Cancer Death",
    rows = outcome == "dod"
  ) %>%
  tab_row_group(
    label = "Distant Metastasis",
    rows = outcome == "mets"
  ) %>%
  tab_spanner(label = "Univariate", columns = c("shr", "shr_lb", "shr_ub")) %>%
  tab_spanner(label = "+ Age", columns = c("shr_age", "shr_lb_age", "shr_ub_age"))

gt_model_full_crr

```

## Heterogeneity Analysis

Here, we enter the estimate and standard error from each model (one for each quartile) into a meta-analysis to determine whether there is heterogeneity between the estimates. Significant heterogeneity between quartiles would indicate that the effect of RP on prostate cancer metastasis/death differs based on a patient's baseline risk.

```{r het}

# Collapse data for meta-analysis
df_meta <-
  left_join(
    df_analysis_crr %>%
      select(outcome, model_results) %>%
      unnest(model_results) %>%
      nest(df_shr = -c(outcome)),
    df_analysis_crr %>%
      select(outcome, model_age_results) %>%
      unnest(model_age_results) %>%
      rename_all(~ paste0(., "_age")) %>%
      rename(outcome = outcome_age) %>%
      nest(df_shr_age = -c(outcome)),
    by = "outcome"
  ) %>%
  mutate(
    # Using the "meta" package, meta-analysis the estimates (subhazard ratios) across quartile groups
    meta_results =
      map(
        df_shr,
        ~ meta::metagen(
          data = ..1,
          TE = coef,
          seTE = se_coef,
          studlab = label,
          sm = "HR",
          comb.fixed = TRUE,
          comb.random = TRUE
        )
      ),
    # Repeat meta-analysis for sensitivity analysis including age
    meta_results_age =
      map(
        df_shr_age,
        ~ meta::metagen(
          data = ..1,
          TE = coef_age,
          seTE = se_coef_age,
          studlab = label_age,
          sm = "HR",
          comb.fixed = TRUE,
          comb.random = TRUE
        )
      )
  )

# Save out information to print in tables

# Distant metastasis outcome
df_meta_results_mets <-
  df_meta$meta_results[df_meta$outcome == "mets"][[1]]

df_meta_results_mets <-
  as_tibble(df_meta_results_mets) %>%
  select(studlab, TE, lower, upper) %>%
  bind_rows(
    tribble(
      ~studlab, ~TE, ~lower, ~upper,
      "Fixed Effects", df_meta_results_mets$TE.fixed,
      df_meta_results_mets$lower.fixed, df_meta_results_mets$upper.fixed,
      "Random Effects", df_meta_results_mets$TE.random,
      df_meta_results_mets$lower.random, df_meta_results_mets$upper.random
    )
  ) %>%
  mutate_at(vars(TE, lower, upper), ~exp(.))

# Prostate cancer death outcome
df_meta_results_dod <-
  df_meta$meta_results[df_meta$outcome == "dod"][[1]]

df_meta_results_dod <-
  as_tibble(df_meta_results_dod) %>%
  select(studlab, TE, lower, upper) %>%
  bind_rows(
    tribble(
      ~studlab, ~TE, ~lower, ~upper,
      "Fixed Effects", df_meta_results_dod$TE.fixed,
      df_meta_results_dod$lower.fixed, df_meta_results_dod$upper.fixed,
      "Random Effects", df_meta_results_dod$TE.random,
      df_meta_results_dod$lower.random, df_meta_results_dod$upper.random
    )
  ) %>%
  mutate_at(vars(TE, lower, upper), ~exp(.))

# Create table with both metastasis and prostate cancer death results
gt_meta_results <-
  left_join(
    df_meta_results_mets %>%
      mutate_at(
        vars(-studlab),
        list("mets" = ~ style_ratio(.))
      ) %>%
      select(studlab, ends_with("_mets")),
    df_meta_results_dod %>%
      mutate_at(
        vars(-studlab),
        list("dod" = ~ style_ratio(.))
      ) %>%
      select(studlab, ends_with("_dod")),
    by = "studlab"
  ) %>%
  mutate(
    ci_mets = str_glue("{lower_mets}, {upper_mets}"),
    ci_dod = str_glue("{lower_dod}, {upper_dod}")
  ) %>%
  filter(!str_detect(studlab, "Effects")) %>%
  flextable::flextable(
    col_keys = c("studlab", "TE_mets", "ci_mets", "TE_dod", "ci_dod")
  ) %>%
  flextable::autofit() %>%
  flextable::set_header_labels(
    studlab = "Quartile",
    TE_mets = "sHR",
    ci_mets = "95% CI",
    TE_dod = "sHR",
    ci_dod = "95% CI"
  ) %>%
  flextable::add_header(
    TE_mets = "Distant Metastasis",
    ci_mets = "Distant Metastasis",
    TE_dod = "Prostate Cancer Death",
    ci_dod = "Prostate Cancer Death",
    top = TRUE
  ) %>%
  flextable::merge_h(part = "header") %>%
  flextable::padding(part = "header", padding.bottom = 5) %>%
  flextable::align(part = "all", align = "center") %>%
  flextable::align(j = "studlab", align = "left")

gt_meta_results

```

# Interaction Analysis

The previous analyses up until this point used baseline risk of BCR as a categorical variable - four groups based on the quartiles of risk distribution in this population. The next analysis tests a similar hypothesis using baseline risk of BCR as a continuous variable, with an interaction term between baseline risk and treatment group (immediate or delayed RP). Again, competing risk regression is used.

```{r interaction}

# Create a dataset which has an interaction between treatment group (rp) and risk of BCR (continuous variable)
df_master_int <-
  df_master %>%
  mutate(int_rp_risk = rp * preopbcr_xb)

# Create for both mets and prostate cancer death
df_analysis_int <-
  tibble(outcome = c("mets", "dod")) %>%
  mutate(
    # Primary analysis
    model_crr_full =
      map(
        outcome,
        ~ cmprsk::crr(
          ftime = df_master_int[[as.character(str_glue("tt{..1}"))]],
          fstatus = df_master_int[[as.character(str_glue("crr_{..1}_outcome"))]],
          cov1 = df_master_int[c("rp", "preopbcr_xb", "int_rp_risk")],
          failcode = 1,
          cencode = 0
        )
      ),
    # Pull out interaction p value
    model_full_int_p =
      map_dbl(
        model_crr_full,
        ~ summary(..1)$coef %>%
          as_tibble(rownames = "var") %>%
          janitor::clean_names() %>%
          filter(var == "int_rp_risk") %>%
          pull(p_value)
      ),
    # Sensitivity analysis including age as a covariate
    model_age_crr_full =
      map(
        outcome,
        ~ cmprsk::crr(
          ftime = df_master_int[[as.character(str_glue("tt{..1}"))]],
          fstatus = df_master_int[[as.character(str_glue("crr_{..1}_outcome"))]],
          cov1 = df_master_int[c("age", "rp", "preopbcr_xb", "int_rp_risk")],
          failcode = 1,
          cencode = 0
        )
      ),
    # Pull out interaction p value
    model_age_full_int_p =
      map_dbl(
        model_age_crr_full,
        ~ summary(..1)$coef %>%
          as_tibble(rownames = "var") %>%
          janitor::clean_names() %>%
          filter(var == "int_rp_risk") %>%
          pull(p_value)
      )
  )

```

# Visualize Interaction

We will then visualize the interaction (if any) using a generalized additive model (GAM). While previously we have been using competing risks regression for all analyses, here the generalized additive model will be based on a Cox model for simplicity. A separate model is created within each treatment group (immediate RP vs delayed RP) and for each outcome (prostate cancer metastasis and death).

As a check on these results, we also created a generalized additive model using logistic regression for the outcome of prostate cancer metastasis or prostate cancer death within 15 years. Patients who did not have metastasis or death within 15 years and did not have at least 15 years of followup were excluded.

```{r visinteraction}

# Need separate model for each outcome and each group
df_analysis_gam <-
  list(
    outcome = c("mets", "dod"),
    rp = c(0, 1),
    outcome_time = c(10, 12, 15)
  ) %>%
  cross_df() %>%
  # TODO: If you don't have 15 years for DOD you can filter it out here
  filter(!(outcome == "dod" & outcome_time == 15)) %>%
  mutate(
    # Save out dataset
    df_model =
      pmap(
        list(rp, outcome, outcome_time),
        ~ df_master %>%
          filter(rp == ..1) %>%
          select(mrn, ends_with("preopbcr_xb"),
                 contains(as.character(str_glue("{..2}"))), -contains("outcome")) %>%
          rename(out := as.character(str_glue("{..2}")),
                 ttout := as.character(str_glue("tt{..2}"))) %>%
          mutate(
            out_timepoint_est =
              case_when(
                out == 1 & ttout <= ..3 ~ 1,
                out == 1 & ttout > ..3 ~ 0,
                out == 0 & ttout >= ..3 ~ 0
              )
          )
      ),
    
    # Create GAM model using Cox regression
    # For Cox GAM models, the outcome needs to be the "time to event" and the actual event indicator goes into the "weights" option
    cox_gam_model =
      map(
        df_model,
        ~ mgcv::gam(
          ttout ~ s(preopbcr_xb, bs = "cs"),
          family = mgcv::cox.ph(),
          data = ..1,
          weights = out,
          method = "REML"
        )
      ),
    
    # Create logistic GAM model for outcome within 10/12/15 years
    logistic_gam_model =
      map(
        df_model,
        ~ mgcv::gam(
          out_timepoint_est ~ s(preopbcr_xb, bs = "cs"),
          family = "binomial",
          data = ..1,
          method = "REML"
        )
      ),
    
    # Get predictions from Cox GAM model
    cox_gam_pred =
      pmap(
        list(cox_gam_model, df_model, outcome_time),
        ~ predict(
          ..1,
          newdata = ..2 %>% mutate(out = 0, ttout = ..3),
          se.fit = TRUE,
          type = "response"
        ) %>% as_tibble(.name_repair = "unique") %>%
          set_names(c(".fitted_coxgam", ".se.fit_coxgam")) %>%
          mutate(.fittedpr_coxgam = 1 - .fitted_coxgam)
      ),
    
    # Get predictions from logistic GAM model
    logistic_gam_pred =
      map2(
        logistic_gam_model, df_model,
        ~ predict(
          ..1,
          newdata = ..2,
          se.fit = TRUE,
          type = "response"
        ) %>% as_tibble(.name_repair = "unique") %>%
          set_names(c(".fitted_loggam", ".se.fit_loggam"))
      ),
    
    # Merge predictions from Cox GAM and logistic GAM models
    df_gam_pred =
      pmap(
        list(df_model, cox_gam_pred, logistic_gam_pred),
        ~ ..1 %>%
          select(preopbcr_xb) %>%
          bind_cols(..2, ..3)
      )
  )

# Save out data for graph
df_gam_final <-
  df_analysis_gam %>%
  select(outcome, rp, outcome_time, df_gam_pred) %>%
  unnest(df_gam_pred) %>%
  select(outcome, rp, contains("preopbcr"), contains("out"),
         .fittedpr_coxgam, .fittedpr_loggam = .fitted_loggam) %>%
  # merge in probabilities that correspond to linear predictors
  # linear predictor used as a covariate in the model but we would like to graph by the actual predicted probability
  left_join(df_master %>% select(preopbcr_xb, preopbcr_pr5), by = c("preopbcr_xb")) %>%
  # Nest separately by outcome and outcome_time
  nest(df_gam_pred = -c(outcome, outcome_time)) %>%
  mutate(
    label =
      case_when(
        outcome == "mets" ~ "Distant Metastasis",
        outcome == "dod" ~ "Prostate Cancer Death"
      )
  )

# Create graph for each outcome
df_gam_graph <-
  df_gam_final %>%
  mutate(
    # Cox GAM model
    gg_cox_gam =
      pmap(
        list(df_gam_pred, label, outcome_time),
        ~ ggplot() +
          geom_density(data = ..1, aes(x = preopbcr_pr5, y = ..scaled..*0.8),
                       col = "gray", fill = "gray", show.legend = FALSE) +
          geom_line(data = ..1 %>% filter(rp == 0), aes(x = preopbcr_pr5, y = .fittedpr_coxgam)) +
          geom_line(data = ..1 %>% filter(rp == 1), aes(x = preopbcr_pr5, y = .fittedpr_coxgam), linetype = "dashed") +
          labs(
            x = "Risk of BCR within 5 Years",
            y = as.character(str_glue("Risk of {..2} within {..3} Years"))
          ) +
          scale_x_continuous(labels = scales::percent_format(accuracy = 1), limits = c(0, 1)) +
          scale_y_continuous(labels = scales::percent_format(accuracy = 1), limits = c(0, 1))
      ),
    
    # Logistic model
    gg_logistic_gam =
      pmap(
        list(df_gam_pred, label, outcome_time),
        ~ ggplot() +
          geom_density(data = ..1, aes(x = preopbcr_pr5, y = ..scaled..*0.8),
                       col = "gray", fill = "gray", show.legend = FALSE) +
          geom_line(data = ..1 %>% filter(rp == 0), aes(x = preopbcr_pr5, y = .fittedpr_loggam)) +
          geom_line(data = ..1 %>% filter(rp == 1), aes(x = preopbcr_pr5, y = .fittedpr_loggam), linetype = "dashed") +
          labs(
            x = "Risk of BCR within 5 Years",
            y = as.character(str_glue("Risk of {..2} within {..3} Years"))
          ) +
          scale_x_continuous(labels = scales::percent_format(accuracy = 1), limits = c(0, 1)) +
          scale_y_continuous(labels = scales::percent_format(accuracy = 1), limits = c(0, 1))
      )
  )

# TODO: This data file will be included in the results file to be sent to be - this should not contain any PHI unless there were any changes made to the code above, so these are just results. I will use this data to graph the results from your cohort along with the results from the other cohorts.

# Save out results from graphs
# TODO: This data will also be saved in the results file to be sent to me - since the graph was created using the df_gam_final data which has no PHI, this will not have PHI either, and will aid me to combining all the graphs for this analysis.
df_gam_graphdata <-
  df_gam_graph %>%
  mutate(
    df_cox_gam =
      map(
        gg_cox_gam,
        ~ ggplot_build(..1)
      ),
    df_logistic_gam =
      map(
        gg_logistic_gam,
        ~ ggplot_build(..1)
      )
  ) %>%
  select(outcome, outcome_time, df_cox_gam, df_logistic_gam)

```

# Bootstrap CI for Difference in Risk

Emily and Andrew discussed creating a figure for the difference in risk between RP and WW groups. To get a confidence interval, we need to bootstrap.

```{r bootstrap}

# Number of bootstrap reps
bootstrapn <- 10

# Create list of bootstrap indicators
# This will be used to determine 
df_bs_map <-
  list(
    boot = c(1:bootstrapn),
    rp = c(0, 1)
  ) %>%
  cross_df() %>%
  mutate(
    bootassign = map(rp, ~ sample.int(nrow(df_master %>% filter(rp == ..1)), replace = TRUE))
  ) %>%
  # Create boot 0 for full cohort - this will give true central estimate
  bind_rows(
    tibble(boot = 0, rp = c(0, 1)) %>%
      mutate(bootassign = map(rp, ~ seq(1, nrow(df_master %>% filter(rp == ..1)), 1)))
  )

# Create list of bootstraps x outcome
df_bs_analysis <-
  # Create a dataset with a separate row for every treatment group/outcome/bootstrap number combination
  list(
    boot = c(0:bootstrapn),
    rp = c(0, 1),
    outcome = c("mets", "dod"),
    outcome_time = c(10, 12, 15)
  ) %>%
  cross_df() %>%
  # TODO: Again, dropping 15 year DOD here if not enough follow-up
  filter(!(outcome == "dod" & outcome_time == 15)) %>%
  left_join(
    df_bs_map,
    by = c("boot", "rp")
  ) %>%
  # Create Cox GAM model for each group
  mutate(
    cox_gam =
      pmap(
        # Pass treatment assignment, list of observations to include in bootstrapped dataset, and outcome (mets or dod)
        list(rp, bootassign, outcome),
        ~ mgcv::gam(
          # Outcome is time to event, predictor is baseline risk
          ttout ~ s(preopbcr_pr5, bs = "cs"),
          family = mgcv::cox.ph(),
          data =
            df_master %>%
            # Keep only treatment group of interest
            filter(rp == ..1) %>%
            # This function keeps those rows which match to the bootstrap assignments
            # This is how we end up with a resampled with replacement dataset
            slice(..2) %>%
            # Rename variables for consistency
            rename(
              out := as.character(str_glue("{..3}")),
              ttout := as.character(str_glue("tt{..3}"))
            ),
          # Cox GAM model syntax puts the outcome variable in the "weights" field
          weights = out,
          method = "REML"
        )
      ),
    
    # Create GAM prediction for 0%-100%
    # This first gives just a prediction ("fit") and standard error ("se.fit")
    cox_pred =
      pmap(
        list(cox_gam, outcome_time),
        # Cox GAM - to get predicted probabilities, give time to outcome, set outcome/event indicator to 0
        # This gives a prediction for every baseline risk from 0% to 100%
        ~ predict(
          ..1,
          newdata = tibble(preopbcr_pr5 = seq(0, 1, length.out = 100), out = 0, ttout = ..2),
          se.fit = TRUE,
          type = "response"
        ) %>% as_tibble()
      ),
    # Merge that prediction and standard error with the actual baseline risk values for graphing
    cox_pred =
      pmap(
        list(cox_pred, outcome_time),
        # Bind column of predicted probabilities to
        ~ bind_cols(
          ..1,
          tibble(preopbcr_pr5 = seq(0, 1, length.out = 100), out = 0, ttout = ..2)
        ) %>%
          # Convert "fit" variable to event probability
          mutate(event_pr = 1 - fit)
      )
  )

# Save out all predictions and calculate difference
df_bs_pred <-
  df_bs_analysis %>%
  select(boot, rp, outcome, outcome_time, cox_pred) %>%
  unnest(cox_pred) %>%
  # Reshape to wide to calculate difference between immediate RP and delayed RP
  select(boot, rp, outcome, outcome_time, preopbcr_pr5, ttout, event_pr) %>%
  pivot_wider(
    names_from = "rp",
    names_prefix = "event_pr",
    values_from = "event_pr"
  ) %>%
  mutate(event_pr_diff = event_pr0 - event_pr1)

# Central estimate is boot = 0, so save this out separately
df_centralest <-
  df_bs_pred %>%
  filter(boot == 0) %>%
  select(-boot)

# To get SE, keep all other bootstrap resamples (except true dataset boot = 0)
df_bs_final <-
  df_bs_pred %>%
  filter(boot != 0) %>%
  # Group by value of preopbcr_pr5
  nest(df_pred = -c(outcome, outcome_time, preopbcr_pr5)) %>%
  # Standard error for bootstrap = standard deviation of difference between groups
  mutate(
    se = map_dbl(df_pred, ~ sd(..1[["event_pr_diff"]], na.rm = TRUE))
  ) %>%
  # Keep 95% CI and merge back in with central estimate
  select(-df_pred) %>%
  left_join(
    df_centralest,
    by = c("outcome", "outcome_time", "preopbcr_pr5")
  ) %>%
  mutate(
    lb = event_pr_diff - 1.96*se,
    lb = if_else(lb < 0, 0, lb),
    ub = event_pr_diff + 1.96*se,
    ub = if_else(ub > 1, 1, ub)
  ) %>%
  select(outcome, outcome_time, preopbcr_pr5, event_pr_diff, lb, ub) %>%
  nest(df_plot = -c(outcome, outcome_time))

# Graph difference between groups with 95% CI                                                                    
df_prdiff_plot <-
  df_bs_final %>%
  # Create label for graph
  mutate(
    label = 
      case_when(
        outcome == "mets" ~ "Distant Metastasis",
        outcome == "dod" ~ "Prostate Cancer Death"
      )
  ) %>%
  # Create graph
  mutate(
    gg_prdiff =
      pmap(
        list(df_plot, label, outcome_time),
        ~ ggplot(..1) +
          geom_line(aes(x = preopbcr_pr5, y = event_pr_diff)) +
          geom_line(aes(x = preopbcr_pr5, y = lb), col = "gray") +
          geom_line(aes(x = preopbcr_pr5, y = ub), col = "gray") +
          labs(
            x = "Risk of BCR within 5 Years",
            y = as.character(str_glue("Risk of {..2} within {..3} Years"))
          ) +
          scale_x_continuous(labels = scales::percent_format(accuracy = 1), limits = c(0, 1)) +
          scale_y_continuous(labels = scales::percent_format(accuracy = 1), limits = c(0, 1))
      )
  )

# TODO: Since this data has no PHI, this will also be included in results sent to me for combining graphs

```

# Discrimination

To calculate discrimination, we will use a Cox model.

```{r discrimination}

# Calculate the c-index for Cox models
df_analysis_disc <-
  tibble(outcome = c("mets", "dod")) %>%
  mutate(
    
    # Data for cox model
    df_model =
      map(
        outcome,
        ~ df_master %>%
          rename(
            out := as.character(str_glue("{..1}")),
            ttout := as.character(str_glue("tt{..1}"))
          )
      ),
    
    # Create cox model
    cox_model =
      map(
        df_model,
        ~ coxph(
          Surv(ttout, out) ~ preopbcr_xb,
          data = ..1
        )
      ),
    
    # Pull out c-index
    cindex =
      map_dbl(
        cox_model,
        ~ ..1$concordance[["concordance"]]
      )
  )

```

# Save Results

```{r save}

# These results are saved for your use when running the report file
save(
  df_master, gt_summary_all, gt_summary_quartiles, Nmets, Nmetscensored, Nmetsdead, Nmetsdod, medianfu,
  gt_model_full_crr, gt_meta_results, df_meta, df_analysis_int, df_gam_graph, df_prdiff_plot, df_analysis_disc,
  file = here("report_results.Rdata")
)


# This is the results file that should be sent to MSK and only contains data points from graphs
save(
  df_gam_final, df_gam_graphdata, df_prdiff_plot,
  file = here("msk_results_to_send.Rdata")
)

```

